---
title: "DGH-GO: Clustering genes disrupted by rare CNVs in ASD patients"
author: "Asif M. "
output: 
    github_document:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

This document shows the utility of DGH-GO in studying the genetic heterogeneity of Autism Spectrum Disorder (ASD). 


* Data set (genes)
  + Rare CNVs disrupting genes in 3802 individuals diagnosed with ASD were collected from Sanders et al. (Sanders SJ, Ercan-Sencicek AG, Hus V, Luo R, Murtha MT, Moreno-De-Luca D, et al. Multiple Recurrent De Novo CNVs, Including Duplications of the 7q11.23 Williams Syndrome Region, Are Strongly Associated with Autism. Neuron. 2011) 

```{r lib, echo=FALSE, include=FALSE}

library(shiny)
library(shinythemes)
library(sparkline)
library(timevis)
library(DT)
library(shinycssloaders)
library(fontawesome)
library(shinydashboard)
library(ggvenn)
library(tidyverse)
library(ggplot2)
library(plotly)
library(shinyWidgets)
library(org.Hs.eg.db)
library(GOSemSim)
library(tidyr)
library(MASS)
library(Rtsne)
library(cluster)
library(vegan)
library(mclust)
library(plyr)
library(clValid)
library("magrittr")
library(factoextra)
library(fpc)
library(dplyr)
library(shinyalert)
library(umap)



```


# Agglomerative hierarchical clustering (AHC)

* AHC clustering was applied to genes semantic similarity matrix
* Rel semantic similarity measure with max combining criterion was employed to input genes
* Biological processes of gene ontology was used to compute similarty scores

```{r ahc_clus, echo=FALSE}

SemSim_mat <- read.csv("E:/Asif_centuri_backups/CenturiWork/clustering/8Feb2022/ApplicationData/SemSim_RareCNVsDisruptedGenes.csv", header = TRUE, sep = ",")


df_hca_in <-SemSim_mat

method_opt_clus <- "silhouette"

#distance matrxi for clustering
rownames(df_hca_in) <- df_hca_in$genes
df_hca_in <- df_hca_in[, -which(names(df_hca_in) %in% c("genes"))]

### PCA
res_pca <- prcomp(df_hca_in, scale = TRUE, rank=50)
fviz_eig(res_pca, ncp = 50)

df_hca_in_pcs <- res_pca$x

df_hca_in <-as.data.frame(df_hca_in_pcs[, 1:9]) 

dist_hca <- dist(df_hca_in)

hca_opt_p <- fviz_nbclust(df_hca_in, 
              hcut, method = method_opt_clus ,
                                    diss =dist_hca, k.max=50)

hca_opt_p <- ggplotly(hca_opt_p)
hca_opt_p

clusters_sil <- hca_opt_p$data


linkage_hca <- "ward"
hca_centers <- 16
res_hca_clus <- agnes(df_hca_in, method = linkage_hca)

class(df_hca_in)

clus <- cutree(res_hca_clus, k = hca_centers)
          
df_hca_in$cluster <-clus 

hca_stat<-cluster.stats(d = dist_hca,  df_hca_in$cluster, alt.clustering = NULL,
                                      noisecluster=FALSE, silhouette = TRUE, G2 = FALSE, 
                                      G3 = FALSE, wgap=TRUE, sepindex=TRUE, 
                                      sepprob=0.1, sepwithnoise=TRUE, compareonly = FALSE, aggregateonly = FALSE)
            

hca_table<-c()
hca_table<-rbind(hca_table,hca_stat$cluster.size, hca_stat$diameter,
                               (hca_stat$average.distance), 
                             hca_stat$median.distance,hca_stat$separation, hca_stat$clus.avg.silwidths
                               , hca_stat$average.between, hca_stat$average.within
                               ,hca_stat$max.diameter, hca_stat$min.separation, hca_stat$avg.silwidth
                               , hca_stat$dunn, hca_stat$dunn2)
            
            
row.names(hca_table)<-c("Cluster size", "Diameter",  "Average distance", 
                                      "Median distance","Separation", "Clus avg. silwidths"
                                      , "Average between","Average within"
                                      ,"Max diameter","Min separation", "Avg. silwidth","Dunn","Dunn2"
            )
colnames(hca_table)<-c(paste("cluster",1:hca_centers,  sep = "."))
            
hca_table<-round(hca_table,3)

# print table


exportOptions = list(
      select = TRUE,
      searching = TRUE, 
      scrollX = TRUE,
      scrollY = TRUE,
      dom = "BRSpfrti",
      buttons = list(
        list(
          extend = "copy",
          text = 'Copy the selected',
          exportOptions = list(modifier = list(selected = TRUE, order    = 'index', page     = 'all', search='none' ))
        ), 
        list(
          extend = "csv",
          text = 'Exported selected to CSV',
          exportOptions = list(modifier = list(selected = TRUE, order    = 'index', page     = 'all', search='none'))
        ),
        list(
          extend = "csv",
          text = 'Exported all to CSV',
          exportOptions = list(modifier = list(selected = FALSE, order    = 'index', page     = 'all', search='none'))
        )
          
      )
    )


# show datatable with exporting options
datatable(
  hca_table,
  caption = "Clustering statistics",
  rownames = TRUE,
  filter = 'top',
  extensions = c("Buttons", "Select"),
  options = exportOptions
    
)


hca_sil <- silhouette(df_hca_in$cluster, dist_hca)
head(df_hca_in)

fviz_silhouette(hca_sil)



dd <- fviz_cluster(list(data = df_hca_in[, -which(names(df_hca_in) %in% c("genes","class", "cluster"))],

                                cluster = df_hca_in$cluster),

                              alpha = 0.9, shape = 19, geom = c("point"), ellipse = FALSE 
                               ) 


hca_plot <- ggplotly(dd)
          
          
          
hca_plot  




g_c_clus <- df_hca_in[, c("cluster")]#"class", "genes",



```

# Clustering after Sil
```{r SilClus, echo=FALSE}

#<= 0 =2; <=0.1=2; 
neg_sil_index <- which(hca_sil[, 'sil_width'] <= 0)

Sil_removed<-df_hca_in[-neg_sil_index,]
dim(Sil_removed)
dim(Sil_removed)
Sil_removed[1:6, 1:10]
  
method_opt_clus <- "silhouette"

#distance matrxi for clustering
sub_Sil_rem <- Sil_removed[, -which(names(Sil_removed) %in% c("cluster"))] #"genes","class", 
dist_hca <- dist(sub_Sil_rem, method = "euclidean")


hca_opt_p <- fviz_nbclust(sub_Sil_rem, #"genes","class",  
              hcut, method = method_opt_clus ,
                                    diss =dist_hca, k.max=50)
hca_opt_p <- ggplotly(hca_opt_p)
hca_opt_p

linkage_hca <- "ward"
hca_centers <- 14

res_hca_clus <- agnes(sub_Sil_rem, method = linkage_hca)

clus <- cutree(res_hca_clus, k = hca_centers)
          
Sil_removed$cluster <-(unname(clus)) 
dim(Sil_removed)

hca_stat<-cluster.stats(d = dist_hca,  Sil_removed$cluster, alt.clustering = NULL,
                                      noisecluster=FALSE, silhouette = TRUE, G2 = FALSE, 
                                      G3 = FALSE, wgap=TRUE, sepindex=TRUE, 
                                      sepprob=0.1, sepwithnoise=TRUE, compareonly = FALSE, aggregateonly = FALSE)
            
hca_table<-c()
hca_table<-rbind(hca_table,hca_stat$cluster.size, hca_stat$diameter,
                               (hca_stat$average.distance), 
                             hca_stat$median.distance,hca_stat$separation, hca_stat$clus.avg.silwidths
                               , hca_stat$average.between, hca_stat$average.within
                               ,hca_stat$max.diameter, hca_stat$min.separation, hca_stat$avg.silwidth
                               , hca_stat$dunn, hca_stat$dunn2)
            
            
row.names(hca_table)<-c("Cluster size", "Diameter",  "Average distance", 
                                      "Median distance","Separation", "Clus avg. silwidths"
                                      , "Average between","Average within"
                                      ,"Max diameter","Min separation", "Avg. silwidth","Dunn","Dunn2"
            )
colnames(hca_table)<-c(paste("cluster",1:hca_centers,  sep = "."))
            
hca_table<-round(hca_table,3)

hca_table






Sil_removed$genes <- rownames(Sil_removed)
g_c_clus_sil <- Sil_removed[, c("genes", "cluster")]


#install.packages("umap")
library(umap)

umap <- umap(Sil_removed[, -c(10,11)],  n_components =4)


df <- data.frame(x = umap$layout[,1],
                 y = umap$layout[,2],
                 z= umap$layout[,3],
                 cluster     = Sil_removed[, 10])

table(df$cluster)
df2 <- df[df$cluster %in% c(3,4,8,10,11),]


p <- plot_ly(x=df2$x, y=df2$y, z=df2$z, type="scatter", mode="markers", color=as.factor(df2$cluster), size=20 )

p <- ggplot(df2, aes(x=x, y=y, colour =as.factor(cluster)     )) +
  geom_point()

p
###########################



```

