---
title: "DGH-GO: Clustering genes to explore the shared etiolgy of complex diseases"
author: "Asif M. "
output: 
    html_document:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

To show another application of DGH-GO in dissecting the shared etiology of complex disorders, genes known for ASD, Developmental Disorders (DD), Intellectual Disability (ID), and Epileptic Encephalopathy (EE) were obtained from Zhang et al.

* Data set
  + source: Zhang Y, Wang R, Liu Z, Jiang S, Du L, Qiu K, et al. Distinct genetic patterns of shared and unique genes across four neurodevelopmental disorders. Am J Med Genet Part B Neuropsychiatr Genet. 2021


```{r SemSim, echo=FALSE, include=FALSE}

library(shiny)
library(shinythemes)
library(sparkline)
library(timevis)
library(DT)
library(shinycssloaders)
library(fontawesome)
library(shinydashboard)
library(ggvenn)
library(tidyverse)
library(ggplot2)
library(plotly)
library(shinyWidgets)
library(org.Hs.eg.db)
library(GOSemSim)
library(tidyr)
library(MASS)
library(Rtsne)
library(cluster)
library(vegan)
library(mclust)
library(plyr)
library(clValid)
library("magrittr")
library(factoextra)
library(fpc)
library(dplyr)
library(shinyalert)
library(umap)

```

# Input genes
* Semantci similarty measure: REL
* Combining criterion: MAX

```{r inputt, echo=FALSE}
semantic_method= "Rel"
semantic_combine="max"

#creating GO object
hsGO <- godata('org.Hs.eg.db', keytype = "SYMBOL", ont="BP", computeIC=TRUE)

dataIn<-read.csv("/home/asif/Desktop/CenturiWork/clustering/8Feb2022/ApplicationData/sharedEtiologyExample_input_code.csv", header = T, sep = ",")

names(dataIn)[1] <- "genes"
table(dataIn$class)
semSimMat <-as.data.frame(mgeneSim(unique(dataIn$genes),
                                       semData=hsGO,
                                       measure=semantic_method,
                                       combine=semantic_combine,
                                       verbose=TRUE))

semSimMat[1:6, 1:6]
semSimMat$genes <- rownames(semSimMat)

semSimMat_withClass <- merge(semSimMat, dataIn, by="genes", all.x=TRUE, all.y=FALSE, sort=FALSE)
# write.csv(semSimMat_withClass, "/home/asif/Desktop/CenturiWork/clustering/8Feb2022/ApplicationData/sharedEtiologySemSimMatrxi_denvo.csv")

```

## Exploratory plots of input

```{r ShowPlots2, echo=FALSE}
 # bar pot
bar_data <- as.data.frame(table(dataIn$class))
colnames(bar_data) <- c("Diseases", "genes")
p <- ggplot( data = bar_data, aes_string( x = colnames(bar_data)[1], 
           y = colnames(bar_data)[2], fill=bar_data[,1]))+ geom_bar(stat="identity")+
  ggtitle("Number of genes per disease set") +
  xlab("Diseases") + ylab("Number of genes") +        coord_flip()

p <-ggplotly(p)

p  
  
#venn diagram
x<-list()

for(i in unique(dataIn$class)){
           set1<- dataIn[dataIn$class== i,][1]
           names(set1) <- i 
           x <- append(x, set1)
       }
  
ggvenn(x, set_name_size = 3)

  
```

# Dimension Reduction (PCA) and Clustering

```{r ahc_clus, echo=FALSE}

# read the semantic similarty matrix
SemSim_mat <-read.csv("E:/Asif_centuri_backups/CenturiWork/clustering/8Feb2022/ApplicationData/sharedEtiologySemSimMatrxi_denvo.csv")

SemSim_mat <- SemSim_mat [-1]
rownames(SemSim_mat) <- SemSim_mat$genes

### PCA
res_pca <- prcomp(SemSim_mat[, -which(names(SemSim_mat) %in% c("genes", "class"))], scale = TRUE, rank=50)
fviz_eig(res_pca, ncp = 50)


df_hca_in_pcs <- res_pca$x
df_hca_in <-as.data.frame(df_hca_in_pcs[, 1:6]) # 
df_hca_in$genes <- SemSim_mat$genes
df_hca_in$class <- SemSim_mat$class

#df_hca_in <- SemSim_mat
dist_hca <- dist(df_hca_in[, -which(names(df_hca_in) %in% c("genes", "class"))])

method_opt_clus <- "silhouette"

#distance matrxi for clustering
hca_opt_p <- fviz_nbclust(df_hca_in[, -which(names(df_hca_in) %in% c("genes", "class"))], 
              hcut, method = method_opt_clus ,
                                    diss =dist_hca, k.max=50 )
hca_opt_p <- ggplotly(hca_opt_p)
hca_opt_p

linkage_hca <- "ward"
hca_centers <- 22
res_hca_clus <- agnes(df_hca_in[, -which(names(df_hca_in) %in% c("genes", "class"))], method = linkage_hca)

clus <- cutree(res_hca_clus, k = hca_centers)
          
df_hca_in$cluster <-(unname(clus)) 

dim(df_hca_in)

hca_stat<-cluster.stats(d = dist_hca,  df_hca_in$cluster, alt.clustering = NULL,
                                      noisecluster=FALSE, silhouette = TRUE, G2 = FALSE, 
                                      G3 = FALSE, wgap=TRUE, sepindex=TRUE, 
                                      sepprob=0.1, sepwithnoise=TRUE, compareonly = FALSE, aggregateonly = FALSE)
            
hca_table<-c()
hca_table<-rbind(hca_table,hca_stat$cluster.size, hca_stat$diameter,
                               (hca_stat$average.distance), 
                             hca_stat$median.distance,hca_stat$separation, hca_stat$clus.avg.silwidths
                               , hca_stat$average.between, hca_stat$average.within
                               ,hca_stat$max.diameter, hca_stat$min.separation, hca_stat$avg.silwidth
                               , hca_stat$dunn, hca_stat$dunn2)
            
            
row.names(hca_table)<-c("Cluster size", "Diameter",  "Average distance", 
                                      "Median distance","Separation", "Clus avg. silwidths"
                                      , "Average between","Average within"
                                      ,"Max diameter","Min separation", "Avg. silwidth","Dunn","Dunn2"
            )
colnames(hca_table)<-c(paste("cluster",1:hca_centers,  sep = "."))
            
hca_table<-round(hca_table,3)
# print table
# write.csv(hca_table, "E:/Asif_centuri_backups/CenturiWork/clustering/manuscript/SharedEtilogy_clusteringStatisticsWith22Clusters.CSV")

exportOptions = list(
      select = TRUE,
      searching = TRUE, 
      scrollX = TRUE,
      scrollY = TRUE,
      dom = "BRSpfrti",
      buttons = list(
        list(
          extend = "copy",
          text = 'Copy the selected',
          exportOptions = list(modifier = list(selected = TRUE, order    = 'index', page     = 'all', search='none' ))
        ), 
        list(
          extend = "csv",
          text = 'Exported selected to CSV',
          exportOptions = list(modifier = list(selected = TRUE, order    = 'index', page     = 'all', search='none'))
        ),
        list(
          extend = "csv",
          text = 'Exported all to CSV',
          exportOptions = list(modifier = list(selected = FALSE, order    = 'index', page     = 'all', search='none'))
        )
          
      )
    )


# show datatable with exporting options
datatable(
  hca_table,
  caption = "Clustering statistics",
  rownames = TRUE,
  filter = 'top',
  extensions = c("Buttons", "Select"),
  options = exportOptions
    
)


            
hca_sil <- silhouette(df_hca_in$cluster, dist_hca)

fviz_silhouette(hca_sil)


plot_df <- df_hca_in[df_hca_in$cluster %in% c(1,2,6,7,8,9,12,13,14,18, 19,20,21,22), ] #18,

dd <- fviz_cluster(list(data = plot_df[, -which(names(plot_df) %in% c("genes","class", "cluster"))],

                                cluster = plot_df$cluster),

                              alpha = 0.9, shape = 19, geom = c("point"), ellipse = FALSE 
                               ) 


hca_plot <- ggplotly(dd)
          
          
          
hca_plot  
############################## UMAP

library(umap)


umap <- umap(plot_df[, -c(7,8,9)],  n_components =2)

df <- data.frame(x = umap$layout[,1],
                 y = umap$layout[,2],
                # z= umap$layout[,3],
                 cluster     = plot_df[, 9])


p <- plot_ly(x=df$x, y=df$y, type="scatter", mode="markers", color=as.factor(df$cluster), size=30 )

p <- ggplot(df, aes(x=x, y=y, colour =as.factor(cluster)     )) +
  geom_point(size=4)

##################################################

g_c_clus <- df_hca_in[, c("genes", "class", "cluster")]

# write.csv(g_c_clus, "E:/Asif_centuri_backups/CenturiWork/clustering/manuscript/SahredEtiology_gens_with_clusters_before_sil_PCAUpdated22Clusters.CSV")

for(i in unique(g_c_clus$cluster)){
  print(i)
  newDf <- g_c_clus[g_c_clus$cluster==i, ]
  print(table(newDf$class))
}


```
